---
title: "NOAA Analysis on Severe Events"
author: "Sebastian Leal"
date: "11/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

Data analysis on U.S. National Oceanic and Atmospheric Administration's
Answering some questions about severe weather events
so that the government or municipal manager can take actions
TODO

## Data Processing

### Download data & uncompress data

Include libraries:
```{r include_libraries, message=FALSE, warning=FALSE}
library(dplyr)
library("ggplot2")
```

Download data:
```{r}
if (!file.exists("noaa_stormdata.csv.bz2")) {
  download.file(
    "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2",
    "noaa_stormdata.csv.bz2"
  )
}
data <- read.csv(bzfile("noaa_stormdata.csv.bz2"))
str(data)

```

Summarize data before plotting it:

```{r}
top15Events <-
  data %>%
  group_by(EVTYPE) %>%
  summarize(
      totalFatalities = sum(FATALITIES, na.rm = TRUE),
      totalInjuries = sum(INJURIES, na.rm = TRUE)
  ) %>%
  arrange(desc(totalFatalities), desc(totalInjuries)) %>%
  top_n(15)
```

In this chart we see the top 15 most harmful events in USA. We see that tornados the most harmful one:

```{r}
ggplot(top15Events) +
  geom_point(color = "red", mapping = aes(x = totalFatalities, y = EVTYPE)) +
  geom_point(color = "blue", mapping = aes(x = totalInjuries, y = EVTYPE)) +
  xlab("RED: Fatalities / BLUE: Injuries") +
  ggtitle("Top 15 most harmful events")

```



## Results

TODO
